---
title: 'Lecture #6'
author: "Nicholas J. Gotelli"
date: "February 04, 2016"
output:
  pdf_document: default
  html_document:
    highlight: tango
    theme: cosmo
---
# Control Structures

## If Statements

- `if` statement
- `if...else` statement
- example of code efficiency versus code readability
- `ifelse` vectorized operations
- `switch` function

# For Loops

- basic structure
- use of numerical index
- exiting early
- cycling over any vector list
- some common mistakes
- for loops versus vectors

Here is some useful code from Stackoverflow:

```{r eval=FALSE}
random.sample <- function(x) {
  success <- FALSE
  while (!success) {
    # do something
    i <- sample(nrow(df), 1)
    x <- df[sample(nrow(df), 1), ]
    # check for success
    success <- x$SCORE > 0
  }
  return(x)
}



```
An alternative is to use `repeat` (syntactic sugar for `while(TRUE))` and `break`:

```{r eval=FALSE}
random.sample <- function(x) {
  repeat {
    # do something
    i <- sample(nrow(df), 1)
    x <- df[sample(nrow(df), 1), ]
    # exit if the condition is met
    if (x$SCORE > 0) break
  }
  return(x)
}
```



\newpage

Function | Group
--------|----------
runif() | 1
rnorm() | 2
rep() | 3
seq() | 4
sort() | 5
order() | 6
which() | 1
sample() | 2
ceiling() | 3
floor() | 4
round() | 5
trunc() | 6
signif() | 1
sort() | 2
table() | 3
unique() | 4
any() | 5
cat() | 6
range() | 1
diff() | 2
cumsum() | 3
cumprod() | 4
cummax() | 5
cummin() | 6
