---
title: 'Bio 264 Homework #2 Solutions'
author: "Nicholas J. Gotelli"
date: "7 February 2016"
output:
  pdf_document: default
  html_document:
    theme: cerulean
---

> As always, there are many ways to answer these questions in R and in Markdown. Study these solutions for any of the problems that gave you trouble. - NJG


Create a new markdown file called `<Your last name>_HW#2_04Feb2016`. Cut and paste the problems below into Markdown, then add chunks of R code to provide your answers. You will turn in just the `.Rmd` file for this homework assignment.

Some of these problems are adapted from:

> Jones, O., R. Maillardet, and A. Robinson. 2009. Scientific Programming and Simulation Using R. CRC Press, Boca Raton.

1. Suppose *x* = 1.1, *a* = 2.2, and *b* = 3.3. First, reproduce each of the following equations in markdown (by setting them within the `$$` boundaries). Next, assign each expression to the value of the variable *z* and print the value stored in *z*.

    a) $x^{a^b}$
    b) $(x^{a})^b$
    c) $3x^3 + 2x^2 + 1$
    d) The digit in the second place of *z* (hint: use `floor()` and/or `&&`)

Here are the equations as you would type them in Markdown:

```
    a) $x^{a^b}$
    b) $(x^{a})^b$
    c) $3x^3 + 2x^2 + 1$
```
And here is the R code for these problems

```{r}

x <- 1.1
a <- 2.2
b <- 3.3

z <- x^(a^b)
print(z)

z <- (x^a)^b
print(z)

z <- 3*x^3 + 2*x^2 + 1
print(z)

```
The last problem (d) is a little tricky. We want the digit in the second place of *z*. At this point, *z* = 7.413, so we are asking R to pull out the digit in the second place, which is a "4". We will start by truncating the decimal piece, then multiplying by 10 and truncating again. That should leave us with the second digit:

```{r}
# Divide z by its integer component and keep the remainder
z1 <- z %% trunc(z)
print(z1)

# Multiply by 10 and then drop the remainder gives answer
z2 <- trunc(z1*10)
print(z2)

# If we want to be compact and do this in one calculation:

trunc((z %% trunc(z))*10)
```

2. Using the `rep` and `seq` functions, create the following vectors:

    a) $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
    b) $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
    c) $(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)$
    
```{r}
c(1:8,7:1)
rep(1:5,1:5)
rep(5:1,1:5)

```

    
3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as *x* and *y* coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R's inverse trigonometry functions (`asin()`, `acos()`, or `atan()`), convert these numbers into polar coordinates (If you don't know what polar coordinates are, read about them on the web or in your calculus textbook).

```{r}


# Create and print the Cartesian coordinates
xyCoors <- runif(2)
print(xyCoors)

# Create a vector to hold the polar coordinates
polarCoors <- vector(mode="numeric",length=2)

# Calculate the vector length
polarCoors[1] <- sqrt(xyCoors[1]^2 + xyCoors[2]^2)

# Calculate the vector angle
polarCoors[2] <- atan(xyCoors[2]/xyCoors[1])

# Print the Polar coordinates
print(polarCoors)
```


4. Suppose that `queue <- c("sheep", "fox", "owl", "ant")` and that `queue` represents the animals that are lined up to enter Noah's Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as 

    a) the serpent arrives;
    b) the sheep enters the ark;
    c) the donkey arrives and talks his way to the front of the line;
    d) the serpent gets impatient and leaves;
    e) the owl gets bored and leaves;
    f) the aphid arrives and the ant invites him to cut in line.
    g) Finally, determine the position of the aphid in the line.
    
```{r}

queue <- c("sheep", "fox", "owl", "ant")

# a) the serpent arrives;
queue <- c(queue,"serpent")
print(queue)

# b) the sheep enters the ark;
queue <-queue[-1]
print(queue)

# c) the donkey arrives and talks his way to the front of the line;

queue <- c("donkey",queue)
print(queue)

# d) the serpent gets impatient and leaves;
queue <- queue[-length(queue)]
print(queue)

# e) the owl gets bored and leaves;
queue <- queue[queue !="owl"]
print(queue)

# f) the aphid arrives and the ant invites him to cut in line.
queue <-c(queue[-length(queue)],c("aphid","ant"))
print(queue)

# g) Finally, determine the position of the aphid in the line.
which(queue=="aphid")
```

5. Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
z <- seq(1,100)

# subset by using the remainder function %% and the union %
z <-z[z %% 2 != 0 & z%%3 !=0 & z%%7 !=0]
print(z)
```

6. Create a vector `z` of 1000 random uniform numbers. 
    a) create a vector that contains 3 numbers: the proportion of the numbers in *z* that are less than 0.10, greater than 0.90, and
    between 0.45 and 0.55.
    
```{r}
z <- runif(1000)
prop <- c(mean(z<0.10),mean(z>0.90),mean(z>0.45 & z<0.55))
print(prop)
```
    
    
b) Making successive copies of *z*, transform your vector of uniform numbers in the following ways:
  - *log* (base 10) of *z*
  - *z^2^*
  - *e^z^* 
  - square root of *z*
c) for each case calculate your vector of 3 numbers to get the new proportions.
     
     
     
d) typeset the formulas from (b) in markdown (with the `$$` brackets). 
   
```{r}

#log10 of z
z <- log10(z)
prop <- c(mean(z<0.10),mean(z>0.90),mean(z>0.45 & z<0.55))
print(prop)

# z^2
z <- z^2
prop <- c(mean(z<0.10),mean(z>0.90),mean(z>0.45 & z<0.55))
print(prop)

# e^z
z <- exp(z)
prop <- c(mean(z<0.10),mean(z>0.90),mean(z>0.45 & z<0.55))
print(prop)

# sqrt(z)
z <- sqrt(z)
prop <- c(mean(z<0.10),mean(z>0.90),mean(z>0.45 & z<0.55))
print(prop)

```

To type set this in Markdown, use:

```
$$ log_{10}(z) $$

$$ z^2 $$

$$ e^z $$

$$ \sqrt{z} $$
```
$$ log_{10}(z) $$

$$ z^2 $$

$$ e^z $$

$$ \sqrt{z} $$
